Title goes here

### 2016-10-23 paper/paper.tex
PAPER: add content

\begin{itemize}
  \item What are the standard routes?
  \item Which flights and airlines diverge from them? In which airports?
  \item How much CO2 would be saved if aircrafts followed a straight-line route
  instead of the standard one?
\end{itemize}

Additionally

\begin{itemize}
  \item How to use raw ADS-B data to identify flights?
\end{itemize}

@@

Our processing algorithm is composed of three phases that are executed in
sequence. The output of a phase is the input of the next. The process starts
from a dataset of raw ADS-B messages encoded in AVRO format, and ends producing
the data necessary to answer our research questions. The phases are the
following:

@@

... The OpenSky network counts 70 active sensors at the time of development,
yielding a pretty good coverage of the western Europe area. Some areas are
covered almost perfectly, and for those we were able to detect flight accurately.

The results from the complete execution of the algorithm are mixed.
It should be considered that there are a lot of minor airports, and flight
information for those airports is, as one could expect, incomplete.
Of 1165 of the total combinations of airports, about 85 \% of them has their
standard route determined by under 10 flights. These pairs usually correspond
to minor airports, and the accuracy of the computed standard route has
necessarily to be considered pretty low. For major airports, however, the
available data rises significantly. As an example, the standard route between
Schiphol and London Gatwick and converse has been determined by 39,
respectively 42 routes.

Tuning the distance function for the route clustering phase was more challanging
than we expected. In many cases, the algorithm ended up considering as diverging
routes that were actually very close to the standard one. We can conclude,
therefore, that even though we think we were able to determine standard routes
with acceptable accuracy and confidence for medium to big airports, diverging
routes detection still has some issues. We do think that the approach of using
sectors and editing distance can give good results in this context, but the
distance function obviously needs much more experiments are fine-tuning to be
able to give good results in all situations.

Observing our results, we conclude that it is definitely possible to
identify flights with acceptable reliability using raw ADS-B data alone. There
are however some issues with the data that make this effort less than
trivial, and some advanced techniques as well as other data sources may be
needed to determine them with more than average accuracy.

>>>>

With this project, we aim to answer the following questions:

\begin{itemize}
  \item What are the standard routes between pairs of airports?
  \item Which flights and airlines diverge from them? In which airports?
  \item How much CO2 would be saved if aircrafts followed a straight-line route
  instead of the standard one?
\end{itemize}

To answer these questions, we are going to use a dataset of raw ADS-B messages
encoded in AVRO format. This dataset provides information from the OpenSky
network for a full week of September 2016. 
The quantity of available messages is very high, but their content is missing
some important informations about the flights, such as airlines and airports. It
follows that there are other questions that must be faced in the development of
the project:

\begin{itemize}
  \item How to use raw ADS-B data to identify flights?
  \item How to use raw ADS-B to discover additional properties of flights, such
  as associated airline or CO2 consumption of the aircraft?
  \item How to identify the departure and arrival airports of a flight?
  \item How to identify standard routes from an unstructured group of flights?
  \item Is ADS-B data alone enough? Do we need external datasets?
\end{itemize}

@@

Our processing algorithm is composed of three phases that are executed in
sequence; the output of a phase is the input of the next. The process starts
from a dataset of raw ADS-B messages encoded in AVRO format, and ends producing
the data necessary to answer our research questions. The phases are the
following:

@@

The OpenSky network counts 70 active sensors at the time of development,
yielding a pretty good coverage of the western Europe area. Some areas are
covered almost perfectly, and for those we were able to detect flight accurately.
Observing our results, we conclude that it is definitely possible to
identify flights with acceptable reliability using raw ADS-B data alone. There
are however some issues with the data that make this effort less than
trivial, and some advanced techniques as well as other data sources may be
needed to determine them with more than average accuracy.

The results from the execution of the route detection algorithm are mixed.
It should be considered that there are a lot of minor airports, and flight
information for those airports is, as one could expect, incomplete.
Of 1165 of the total combinations of airports, about 85 \% of them has their
standard route determined by under 10 flights. These pairs usually correspond
to minor airports, and the accuracy of the computed standard route has
necessarily to be considered pretty low. For major airports, however, the
available data rises significantly. As an example, the standard route between
Schiphol and London Gatwick and converse has been determined by 39,
respectively 42 routes.

Tuning the distance function for the route clustering phase was more challanging
than we expected. In some cases, the algorithm ended up considering as diverging
routes that were relatively close to the standard one. We can conclude,
therefore, that even though we think we were able to determine standard routes
with acceptable accuracy and confidence for medium- to big-size airports,
diverging routes detection still has some issues. We do think that the approach
of using sectors and editing distance can give good results in this context, but
the distance function obviously needs more experiments are fine-tuning to
be able to give good results in all situations.
